---
title: "Section 8.2 Problems"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    highlight: "kate"
    df_print: "kable"
    includes:
      in_header: "notes_text_preamble"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)

library(tidyverse)
```

## 1.

\[
  \begin{bmatrix}
   6 & - 7/2\\
   -7/2 & 8
  \end{bmatrix}
\]


## 3.

\[
  \begin{bmatrix}
    3 & 0 & 3\\
    0 & 4 & 7/2\\
    3 & 7/2 & 7
  \end{bmatrix}
\]


## 4.

Positive definite.

## 5.

Negative definite.


## 7.

Indefinite.

## 8.

$A^2$ is positive semidefinite because all its eigenvalues are 0 or positive. IF $A$ is positive definite, so is $A^2$ by that same token.

## 9.

### a.

Symmetric. $(A^T)^2 = (-A)^2 = A^2$.

### b.

It is negative semi-definite because the eigenvalues of $A$ are either
0 or purely imaginary:

\[\begin{aligned}
  AX  = \lambda x\\
  &x^HA^H = \bar \lambda x^H\\
  & -x^HA = \bar \lambda x^H\\
  & - \bar \lambda x^H = \bar \lambda x^H\\
  & - \bar \lambda = \bar \lambda
  \end{aligned}
\]
so the eigenvalues of $A^2$ are either 0 or real positive.

### c.

None, since there are imaginary eigenvalues.

## 10.

\[ \begin{aligned}
  & T(X) = (x^TA + v^TA)(x+v) - x^TAx - v^Tv\\
  & = x^tAx + v^Tax + x^TAv + v^TAv - x^TAx - v^TAv\\
  & = v^TAx + x^TAv\\
  & = 2 x^TAv
  \end{aligned}
\]

## 11.

  The same - the eigenvalues don't change sign.

## 12.


$\det A <0$ if and only if
the eigenvalues have opposite signs. Thus $\lambda_1c_1^2 + \lambda_2c_2^2$ is indefinite
because of the differing signs.

## 13.

Let $d$ be a a diagonal emenet less than or equal to zero. Then
the term $dx_i^2$ is zero or negative definite, and so the matrix
cannot be positive definite.

## 14.

Since $\det A >0$ and $a >0$, then
if $c \leq 0$ $ac \leq b^2$, which
violates the constraint. And if $\det A > 0$ then all eigenvalues are positive, so $A$ is positive definite.


## 23.

No, since $M$ is not symmetric.

## 24.

$a$.

## 25.

$\lambda$, since the unit eigenvectors eigenbasis coordiantes are 1.

## 26.

True. Such a vector satisifeis $Ax = \lambda x$ for $\lambda = 0$. Then$A$ has zero determinant and is
therefore not invertible.

## 27.

A hyperellipse whose axes are the
unit eigenvectors.

## 30.

$A =  (Q \sqrt \Lambda Q^T)^2$

If  amatrix is positive definite, so is its square root, since the eigenvalues remain positive.


## 32.

for the $2 \times 2$

Solving the system gets su


\[
\begin{aligned}
& x  = \sqrt a\\
& y = b /\sqrt a\\
& z = \sqrt {c - b^2/a}
\end{aligned}
\]

## 33.

\[
  L = \begin{bmatrix}
  \sqrt 8 & -2/ \sqrt 8\\
  & -2 / \sqrt 8 & \sqrt 5 + 1/2
  \end{bmatrix}
\]

## 34.


If $A$ is positive definite, then it sfull quadratic form is
positive definie. For a $3 \times 3$, the form of the first submatrix is $ax_1^2$. This must be positive definite, because if not then the
form of the second leading submatrix $ax_1^2 + bx_1x_2 +cx_2^2$ would
not be positive definite, and so forth.

And if all the elading submatrices have psoitive definite, then
they have all positive eigenvalues, meaning they have positive determinant.


Then for the final decomposition, to find $x$ we ahve to solve the system:

\[
\begin{aligned}
& a_n1 = x_1 b_{11}\\
& a_n{2} = x_1b_{12} +x_2 b_{22}\\
\vdots
& a_{n{n-1}} = x^T b{n-1}
\end{aligned}
\]

which can clearly be solved in sequence,
with the last element of $x$ ignored because $B$ is triangular with dimension $n-1$. The final equation is

\[
  a_{nn} = x^Tx +t
\]
and $t$ must be positive to split it into $\sqrt t$ across $L$ and $L^T$.


## 36.


```{r}
to_solve <- cholesky <- matrix(0, nrow = 3, ncol = 3)
A <- matrix(c(4, -4, 8, -4, 13, 1, 8, 1, 26), nrow = 3)
u <- sqrt(A[1, 1])
v <- A[1, 2] / u
w <- sqrt(A[2, 2] - v^2)
x <- A[3, 1] / u
y <- (A[2, 3] - x * v) / w
z <- sqrt(A[3, 3] - x^2 - y^2)
to_solve[lower.tri(to_solve, diag = TRUE)] <- letters[seq(length(letters) - 6), lenght(letters)]
cholesky[lower.tri(cholesky, diag = TRUE)] <- c(u, v, x, w, y, z)
print_eqn(mat2latex(to_solve), mat2latex(t(to_solve)), " = ", mat2latex(A))
print_eqn(mat2latex(L), " = ", mat2latex(cholesky))
```

## 36.

\[
    A = LL^T = QR \implies L = QRL^T{-1}
\]

## 37. 

These two criteria are another way of saying that the derivative matrix is positive definite. In that case, the 
product of the separate derivatives is greater than the derivative of the products, which is a condition of the test for a global minimum

## 38.

$p-q > 0$.


## 40.

If $\lambda_k$ has the greatest absolute value of the negative-signed eigenvalues, $k > \lambda_k$ does the job.

## 41.

3.

## 42.

$\frac{n(n+1)}{n}$

## 43.

Yes. Image is degree-0 polynomials, kernel all matrices with $a=0$,
rank 1, nullity 3.

## 44.

Yes, isomorphic since it preserves all three coordiantes.

\[
    \begin{bmatrix}
    a & b & c\\
    b & d & e\\
    c & e & f
    \end{bmatrix}
\]

## 45.

$T(x) = (a+2c+ +2e +f)x_1^2 + (b + d+ e)x_2^2$
Image: all degree-1 polynomials
Kernel: quadratic forms where only $x_3^2$ has a nonzero coefficient
Rank 2, nullity 4.




## 46.

$T(x) = 4ax_1^2 +x_2^2 + 4x_1x_2$.
Image: al qaudratic forms 


## 52.

Consider a $3\times 3$.

\[
    a = e_1^t a_1 e = 1(1)a = a
\]
and

\[
\begin{align}
    b = 1/2((e_1 + e_2)^TA(e_1 +e_2)^T  - e_1^TAe_1 - e_2^TAe_2\)\
    & = 1/2(e_1^TAe_1 + e_2^TAe_2 + e_1^TAe_2 + e_2^TAe_1 - e_1^TAe_1 - e_2^TAe_2)\\
    & = 1/2(e_1^T Ae_2 + e_2^TAe_1)\\
    & = 1/2(b + b)\\
    & = b
\end{align}
\]

## 51.

Alternating odd, even.


## 53.

### a.

$p(x) = x^2 a_{ii} + 2xy a_{ij} +y^2 a_{jj}$

### b.

If so, then $a_{ii}$ and $a_{jj}$ are positive and 
$a_{ii}a_{jj} > a_{ij}^2$, which mean the matrix of
$p$ ahs positive eigenvalues and is positive definite.

### c.

Then $2_{aji}=a_{jj}a_{ii}$, which translates to a zeroeigenvalue and an indeterminate matrix.


## 54.

They must be 0; otherwise the final determinant is negative, so
some eigenvalues are negative and the matrix is not positive semidefinite.

## 55.

If this were not true the determinant would be negative, preventing he matrix from being positive definite.

## 56.

A symmetric $A$ may be factorized $A = Q \Lambda Q^T$, where $Q$ is the orthogonal eigenvectors matrix. Then $a_11  = \lambda_1 x_{11}^2 + \dots + \lambda_n x_{1n^2$. $x_i^2 \leq 1$ since the se are unit vectors. If so, then $a_11 = 1 \lambda_i = \lambda_i$. This is a quadratic form with the eigenvalues as the cosntants. So $a$ is maximally $\lambda_{max}$. Since $a= \Lambda q^T_1 \codt q^T_1$, by Cauchy-Swarz

\[\begin{align}
    &| \Lambda q^T_1 \cdot q^T_1| \leq ||\Lambda 1^T_1 \cdot q^T_1|| \cdot ||q^T_1||
    & \leq ||\Lambda 1^T_1 \cdot q^T_1||\\
    & \leq \lambda_{max}
\end{align}
\]
so $a$ cannot be greater than the greatest $\lambda_i$


## 58.

A spheroid around the origin.

## 58.

An ellipse around the origin

## 59.

A line.

## 60.

Convcave surfaces oreitned positively.

## 61.

Concave surfaces oriented negatively.

## 63.

The expression is eigenbasis coordinates divided by the appropriate lambdas, which cancel.
\[
    q(c_1w_1 + \dots + c_nw_n) = \lambda_1 c_1^2/\lambda_1 + \dots + \lambda_n c_n^2 / \lambda_n = c_1^2 + \dots + c_n^2 
\]

## 64.

```{r}
eigens <- eigen(matrix(c(8, -2, -2, 5), nrow = 2))
eigens$vectors / sqrt(eigens$values)
```

## 65.

For the negative-signed $\lambda$

\[
    c_i^2 \frac{v_i}{\sqrt {|\lambda_i}}^2 = \lambda_i c_i^2 / |lambda_i| = - c_i^2
\]

## 66.


```{r}
eigens <- eigen(matrix(c(3, -5, -5, 3), nrow = 2))
eigens$vectors[, order(eigens$values, decreasing = TRUE)] * 1 / sqrt(abs(sort(eigens$values, decreasing = TRUE)))
```

## 67. 

By the above logic, if we scale the eigenvectors with negative-signed eigenvalues by $1/ \sqrt{|\lambda_i}$, $c_i^2$ is opposite-signed.


## 68.


$R^TAR$.

## 69.

$R$ has rank $m$.
