---
title: "Section 5.3 Exercises"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.pos = "", warning = FALSE,
                      tidy = TRUE, fig.align = "center")

library(tidyverse)
library(rlang)
library(matador)
```


## 1.

$F_{k}$ = $F_{k-1} + F_{k-2}$. So $F_k$ is even if both the preceding two numbers are both even and odd, odd otherwise. Given the starting sequence $0, 1$, $F_3$ is odd, $F_4$  is even, $F_5$ is odd, as is $F_6$, and $F_7$ is again even. The pattern repeats indefinitely.

## 2.

```{r, results='asis'}
m <- square(0, .5, 0, 0, 0, 1/3, 6, 0, 0)
lambdas <- eigen(m, only.values = TRUE)

mat2latex(m %^% 3)
```


b.Each iteration represents 1 year, so after two three-year cycles there will be 3000 again.

## 3.

```{r}
mat_pows(square(1, 1, 1, 0), 2:4)
```


##4.

The "Gibonacci" matrix is:
\[
\begin{bmatrix}
  1/2 & 1/2\\
  1 & 0
\end{bmatrix}
\]

\[
  \begin{aligned}
    & \lambda^2  + 1/2 \lambda - 1/2 = 0
    & \lambda = (1, -1/2)\\
    & S = \begin{bmatrix}
      1 & -2\\
      1 & 1
    \end{bmatrix}
  \end{aligned}
\]

b. The limit is 

\[ \begin{bmatrix}
  2/3 & 1/3\\
  2/3 & 1/3
\end{bmatrix}\]

c.
```{r, results='asis'}

S <- square(1, 1, -1/2, 1)
C <- solve(S) %*% c(1, 0)

mat2latex(S %*% diag(x = c(1, -1/2)^100) %*% C)
```

## 5.

\[S^{-1} = \frac{1}{\lambda_1 -\lambda_2}\begin{bmatrix}
  1 & -\lambda_2\\
  -1 & \lambda_1
\end{bmatrix}\]

\[
  \begin{aligned}
    & c = \begin{bmatrix}
  1 & -\lambda_2\\
  -1 & \lambda_1
\end{bmatrix} \begin{bmatrix}
  1\\
  0
\end{bmatrix} = \begin{bmatrix}
  \frac{1}{l\lambda_1 - \lambda_2}\\
  -\frac{1}{l\lambda_1 - \lambda_2}
\end{bmatrix}\\
& F_k = \begin{bmatrix}
  \lambda_1 & \lambda_2\\
  1 & 1
\end{bmatrix} 
\begin{bmatrix}
  \lambda_1^k & 0\\
  0 & \lambda_2^k
\end{bmatrix}
\begin{bmatrix}
  \frac{1}{l\lambda_1 - \lambda_2}\\
  -\frac{1}{l\lambda_1 - \lambda_2}
\end{bmatrix}\\
& = \begin{bmatrix}
  \lambda_1\lambda_1^k & \lambda_2\lambda_1^k\\
  \lambda_1^k & \lambda_2^k
\end{bmatrix}
\begin{bmatrix}
  \frac{1}{l\lambda_1 - \lambda_2}\\
  -\frac{1}{l\lambda_1 - \lambda_2}
\end{bmatrix}\\
&F_k= \frac{\lambda_1^k - \lambda_2^k}{\lambda_1 - \lambda_2}
\end{aligned}
\]

## 7.

The Lucas matrix is a Fibonacci variant that starts with $\m{1\\2}$

The sequence goes $2, 1, 3, 4, 7, 11, 18, 39, 57, 76, 133$. And $\frac{1}{2}(1 - \sqrt5)^10 = 133$.

## 8.

$\lambda = 1, 3/4, 1/2$, so the steady state is $\m{1\\0\\0}$ - all dead.

##9.

\[
 \begin{bmatrix}
   5/6 & 1/6 & 0\\
   1/6 & 5/6 & 0\\
   1/4 & 1/3 & 1
 \end{bmatrix}
\]

## 10.

The provided matrix _is_ $A$, so we diagonalize it and multiply by $u_0$ to 
complete $u_k=A_ku_0$
The eigenvector corresponding to $\lambda = 1$ is $\m{3/2\\1}$, so that is the equilibrium.$

```{r}
A <- square(.8, .2, .3, .7)
eigen(A)
```

\[
  \begin{aligned}
    & u_k = \frac{2}{5} \begin{bmatrix}
      3/2 & -1\\
      1 & 1
    \end{bmatrix}
    \begin{bmatrix}
      1^k & 0\\
      0 & .5^k
    \end{bmatrix}
    \begin{bmatrix}
      1 & 1\\
      -1 & 3/2
    \end{bmatrix}
    \begin{bmatrix}
      0\\
      5
    \end{bmatrix}\\
    & = \frac{2}{5} \begin{bmatrix}
      3/2 & -(.5^k)\\
      1 & .5^k
    \end{bmatrix}
    \begin{bmatrix}
      5\\
      15/2
    \end{bmatrix}\\
    & = \frac{2}{5}
    \begin{bmatrix}
    15/2 - \frac{15(-.5k)}{2}\\
    5 + \frac{15(.5^k)}{2}
    \end{bmatrix}\\
    & =\begin{bmatrix}
      5 - 5(.5)^k\\
      2 + 5(.5)^k
    \end{bmatrix}
  \end{aligned}
\]

## 11.

a. $v_1 + v_2 =2v_3$, so for $\lambda =0$ the eigenvector is $x = \m{-1/2\\-1/2\\1}$. So the

I cheat a little by borrowing a function from StackOverflow to compute the standard eigenvectors.
I use it to find the formula.
```{r}
# From https://stackoverflow.com/questions/43223579/solve-homogenous-system-ax-0-for-any-m-n-matrix-a-in-r-find-null-space-basi/43225005#43225005
NullSpace <- function (A) {
  m <- dim(A)[1]; n <- dim(A)[2]
  ## QR factorization and rank detection
  QR <- base::qr.default(A)
  r <- QR$rank
  ## cases 2 to 4
  if ((r < min(m, n)) || (m < n)) {
    R <- QR$qr[1:r, , drop = FALSE]
    P <- QR$pivot
    F <- R[, (r + 1):n, drop = FALSE]
    I <- base::diag(1, n - r)
    B <- -1.0 * base::backsolve(R, F, r)
    Y <- base::rbind(B, I)
    X <- Y[base::order(P), , drop = FALSE]
    return(X)
    }
  ## case 1
  return(base::matrix(0, n, 1))
}
A <- square(.2, .4, .3, .4, .2, .3, rep(.4, 3), byrow = TRUE)

S <- sapply(eigen(A)$values, function(x) NullSpace(A - diag(x= x, nrow = 3)))
S
```

b., c.

\[ =\begin{bmatrix}
  .75 & -1 & -.5\\
  .75 & 1 & -.5\\
  1 & 0 & 1
\end{bmatrix} \begin{bmatrix}
  1^k & 0 & 0\\
  0 & -.2^k & 0\\
  0 & 0 & 0
\end{bmatrix}
\begin{bmatrix}
  .4 & .4 & .4\\
  -.5 & .5 & 0\\
  -.4 & -.4 & .6
\end{bmatrix}
\begin{bmatrix}
  0\\
  10\\
  0
\end{bmatrix}\\
= \begin{bmatrix}
  .75(1^k) & -1(-.2^k) & 0\\
  .75(1^k) & -.2^k & 0\\
  1^k & 0 & 0
\end{bmatrix}
\begin{bmatrix}
  0\\
  5\\
  0
\end{bmatrix}\\
= 5(-.2^k)\]

```{r}
NullSpace(A)
```

## 12.


The matrix is 

\[
  A = \begin{bmatrix}
    1/2 & 0 & 1/2\\
    0 & 1/2 & 1/2\\
    1/2 & 1/2 & 0
  \end{bmatrix}
\]

For $\lambda = 1$, $x = \m{1\\1\\1}$

\[
  \begin{bmatrix}
    -1/2 & 0 & 1/2\\
    0 & 1/2 & 1/2\\
    1/2 & 1/2 & 0
  \end{bmatrix}
\]

Scale to unit length for 

\[U_{\infty} = \begin{bmatrix}
  1/\sqrt3\\
  1/\sqrt3\\
  1/\sqrt3
\end{bmatrix}\]

## 13.

a. $0 \leq a,b \leq 1$.

b. The eigenvalues are 1 and $a-b$ (since the trace is $1 + a-b$)

The eigenvectors:
\[
\begin{aligned}
& S =\begin{bmatrix}
  \frac{1-a}{b} & -1\\
  1 & 1
\end{bmatrix}
\\ & S^{-1} = \Big(\frac{b}{1-a} +1 \Big) \begin{bmatrix}
  1 & 1\\
  -1 & \frac{1-a}{b}
  \end{bmatrix}\\
& \Lambda = \begin{bmatrix}
  (1 -a)^k & 0\\
  0 & 1^k
\end{bmatrix}
\end{aligned}
\]

\[
  \begin{aligned}
    & c = S^{-1} = \Big(\frac{b}{1-a} +1 \Big) \begin{bmatrix}
  1 & 1\\
  -1 & \frac{1-a}{b}
  \end{bmatrix}
  \begin{bmatrix}
    1\\
    1
  \end{bmatrix}\\
  & =
  \begin{bmatrix}
    2(1 + \frac{b}{1-a}\\
    0
  \end{bmatrix} \\
  & \lambda c_1 x_1 = 2( 1 + \frac{b}{1-a})
  \begin{bmatrix}
    \frac{1-a}{b}\\
    1
  \end{bmatrix}\\
  & = \begin{bmatrix}
    4\\
    2( 1 + \frac{b}{1-a})
  \end{bmatrix}
  \end{aligned}
\]

I think it's $\m {2/3\\1/2}$

## 14.

a. For the stable state:
```{r}
A <- square(rep(.5, 3), .25, .5, 0, .25, 0, .5, byrow = TRUE)

S <- eigen(A)$vectors
Lambda <- diag(eigen(A)$values)

stable <- S[,1]
MASS::fractions(stable / sum(stable))
```
The diagonalization is:

b.
```{r, results = "asis"}
S <- map(eigen(A)$values, ~NullSpace(A - diag(nrow = 3, x = .x))) %>% 
  reduce(cbind)
invisible(sapply(list(S, Lambda,solve(S) ), mat2latex))
```

## 15.

Say $m =3$ and $S = \m{1 & 1 & 1}$.
\[
  \begin{aligned}
    & SA = S\\
    & S x = \sum^{m}_{i=1}{x_i}\\
    & SAx = Sx = \sum^{m}_{i=1}{x_i}
  \end{aligned}
\]

## 18.

## 19.

\[
  \begin{aligned}
    & I = (I -A)(I + A + A^2 + \dots + A^n)\\
    & = (I^2 - A) + (A - A^2) + (A^2 - A^3) +(A^n - A ^{n+1})\\
    & = I
  \end{aligned}
\]

This works when no eigenvalue is greater than 1; the sequence has a finite
sum, so $(A-I)$ has an exact inverse.
```{r, results = "asis"}
A <- diag(x =0, nrow = 3)
A[upper.tri(A)] <- 1

mat2latex((diag(x =3) - A) %*% (mat_pows(A, 1:100) %>% reduce(`+`) + diag(x =3)), sink = FALSE)
```

## 20.


```{r}
A <- square(0, 0, .2, .5)
A %^% 100
```

\[
  \begin{aligned}
  & A^k = \frac{1}{6} \begin{bmatrix}
    1 & .4\\
    0 & 1
  \end{bmatrix}
  \begin{bmatrix}
    0 & 0\\
    0 & .5^k
  \end{bmatrix}
  \begin{bmatrix}
    1 & -.4\\
    0 & 1
  \end{bmatrix}
  \end{aligned}
\]

This does sum to `r mat2latex(solve(diag(nrow = 2) - A), sink = FALSE) `

## 24.

\[B = \begin{bmatrix}
  3 & 1\\
  0 & 2
\end{bmatrix}\]

\[B = \begin{bmatrix}
  1 & -1\\
  0 & 1
\end{bmatrix}
\begin{bmatrix}
  3^k & 0\\
  0 & 2^k
\end{bmatrix}
\begin{bmatrix}
  1 & 1\\
  0 & 1
\end{bmatrix}
\]

## 26. 

Identical.

## 28.

a. The eigenvectors for $\lambda =0$ always span the kernel, since they are solutions for $Ax =0$.

b. When they are real.

## 29. 

If $B^4= I$ and $C^3 = -I$, they are rotations about $\pi/2$ and $\pi/3$ radians respectively.

Since an angle on the complex plane is is $\cos \theta i \sin \theta$ + 
\[
  \begin{aligned}
    & \lambda = e^{\pm i(\pi/2)} = \pm i
    & \lambda = e^{i (\pi/3)} = \frac{1 \pm \sqrt{-3}}{2} 
  \end{aligned}
\]
