---
title: "Section 5.2 Problems"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.pos = "", warning = FALSE,
                      tidy = TRUE, fig.align = "center")

library(tidyverse)
library(rlang)
library(matador)
```

## 7.
Got it in one!

```{r, results = "asis"}
v_1 <- c(2, 2, 1)
v_2 <- c(-2, 1, 2)
v_3 <- c(18,0,0)

u_1 <- normalize(v_1)
v_2p <- v_2 - (t(v_2) %*% u_1) * u_1
u_2 <- normalize(v_2p)
v_3p <- v_3 -(t(v_3) %*% u_2)*u_2 - (t(v_3) %*% u_1) *u_1
u_3 <- normalize(v_3p)

R <- diag(x = apply(cbind(v_1, v_2p, v_3p), 2, l2) )
R[1,2] <- t(v_2) %*% u_1
R[1,3] <- t(v_3) %*% u_1
R[2, 3] <- t(v_3) %*% u_2
print_eqn(c("R =", mat2latex(R, sink = TRUE)))
Q <- cbind(u_1, u_2, u_3)
print_eqn(c("Q =", mat2latex(round(Q,  3), sink = TRUE)))
mat2latex(Q %*% R)
```

## 11.


\[A = \begin{bmatrix}
4/5 &-1/5\\
0 & 2/15\\
0 & 14/15\\
3/5 & 4/15
\end{bmatrix}
\begin{bmatrix}
5 & 10\\
0 & 15
\end{bmatrix}
\]
## 31.

The QR on the generic matrix
\[\m{a & b &d\\
0 &c & e \\
0 & 0 &f}\]

is just

\[\m{1&0 &0\\
0&1&0\\
0 &0 &1}
\m{a & b &d\\
0 &c & e \\
0 & 0 &f}\]

A convenient property of upper-triangular matrices,
since only one dimension of vector $j$ is left after subtracting $u_{j-1}$.

## 32.

A basis of $x_1 + x_2 +x_3 = 0$ is $\m{-1/ \sqrt 2& -1 / \sqrt 2\\ 1 & 0\\ 0 & 1}$.

## 36.

$M$ is already in QR form.

## 37.

Just $A$ with the columns normalized and $R$ with lengths in the diagonal,
since the vectors are already orthogonal.

## 39.

We can construct an orthonormal basis given some
vectors by finding a normal vector from the kernel
of the transpose.

```{r, results = "asis"}
u_1 <- normalize(1:3)
u_2 <- normalize(c(1, 1, -1))
u_3 <- normalize(c(5, -4, 1))

mat2latex(pracma::rref(t(cbind(u_1, u_2))))

mat2latex(normalize(c(5, -4, 1)))
```

## 40.

If a matrix has orthogonal but not orthonormal columns, then $Q$ is each vector scaled to unit length and $R$ is a diagonal matrix with diagonal $||v_j||$, since none of the vectors has any component parallel to another.

## 41.

As I showed above, the $QR$ of an upper triangular matrix consists of the standard basis and the original matrix

## 44.

If $A$ lacks full column rank, then $v^{\perp}$ for redundant columns is 0, because they are linear combinations of the others. So column $j$ of $Q$ is 0 and element $jj$ of $R$ is 0 as well, since the zero vector has no length. That means the matrix is not invertible, meaning there is no unique factorization.

## 45.

In this case, $v^{\perp}$ is always nonzero because the vectors are all linearly independent, so none point in the same direction. By induction, $v_2^{\perp}$ is nonzero because $v_2$ is not a multiple of $v_1$, and $v_3^{\perp}$ is nonzero because no constants satisfy $c_1v_1 +c_2v_2 = v_3$. If $R$ is instead lower triangular $L$, then
the diagonal remains the lengths of each $v^{\perp}$, and th lower triangle
reflects the upper triangle of $R$. In other words, $L=R^T$.
\[
\begin{bmatrix}
  0 & 2 & 3\\
  4 & 5 & 6
\end{bmatrix}
\]
