---
title: "Section 4.2 Problems"
author: "Ryan Heslin"
date: "`r Sys.Date()`"
output: pdf_document
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.pos = "", warning = FALSE,
                      tidy = TRUE, fig.align = "center")

library(tidyverse)
library(rlang)
library(matador)
```

We check the linearity of some transformations.

## 1.

$T(M) = M + I_2$: nonlinear

## 3.

\[T(M) = \text{trace}(M)\]
Obviously not isomorphic; many matrices have the same trace. 

##5. 

$T(m) = M^2$: not isomorphic

## 7.

\[T(M) = 
  \begin{bmatrix}
  1 &2 \\ 3 &4
  \end{bmatrix}M
\]

Obviously linear, given matrix \[
  \begin{bmatrix}
  1 & 0 & 1 &0\\
  3 &0 & 4 & 0\\
  0 & 1 & 0 &1\\
  0 & 3 & 0 &4
  \end{bmatrix}
\]

Also an isomorphism, since this is clearly invertible.

## 9.

\[ T(M) =
  \begin{bmatrix}
  3 & 4\\
  5 & 6
  \end{bmatrix}^{-1}
  M \begin{bmatrix}
  3 & 4\\
  5 & 6
  \end{bmatrix}
\]

Not an isomorphism, since $M$ itself may be noninvertible.

Actually yes:

\[
  \begin{aligned}
    & T(M) = S^{-1}MS\\
    & S^{-1}M = S^{-1}\\
    &M = SAS^{-1}
  \end{aligned}
\]

## 12.

\[T(c) =cM \quad c \rightarrow R^2\]
where $M$ is a particular matrix.
Linear -  $c(M_1+M_2)=cM_1+cM_2$ and $c(kM_1)=kcM_1$.

Also isomorphic: $M = 1/c T(c)$

## 16.

\[T(M) = M \begin{bmatrix}
  2 & 0\\
  0 & 3
\end{bmatrix} - \begin{bmatrix}
  3 & 0\\
  0 & 4
\end{bmatrix}M\]

The matrix is:

\[\begin{bmatrix}
  -1 & 0 & 0 & 3\\
  0 & 0 & 0 & 3\\
  0 & 0 & 0 & 0\\
  -2 & 0 & 0 & -1
\end{bmatrix}\]

Linear, not invertible.

## 17.

\[T(x + iy) = x\]

Linear, not invertible, as the complex part is just zeroed out.
## 18.

$T(x + iy) = x^2 +y^2$

Non-linear ($cT(x) = cx^2 +cy^2$), but $T(cx) =c^2x^2 +c^2y^2)$; $x$ can be positive or negative.

## 21.

\[T(x + iy) = y+ix\]

Linear: $T(x_1+iy_1 + x_2 +iy_2) = y_1 + y_2 +i(x_1 + x_2) = T(x_1+iy_1) + T(x_2 +iy_2)$ and $T(kx_1 +kiy_1) = ky_1 + kx_1i_y = kT(x+iy)$

Isomorphism, and involutory too: $T^{-1}y= y +ix$.

## 22.

\[T(f(t)) = \int_-2^3f(t)dt\]
\[
\begin{bmatrix}
  3 & 3/2 & 1
\end{bmatrix} -
\begin{bmatrix}
  2 & 1 & 2/3
\end{bmatrix}
\]

Clearly noninvertible; infinitely many functions with the same definite integral.
## 23.

\[T(f(t)) = f(7)\]
\[T(f(t)) = a + 7b + 49c\]

Linear, nonisomorphic, with matrix 
\[
  \begin{bmatrix}
  1 & 7 & 49\\
  0 & 0 & 0\\
  0 & 0 &0
  \end{bmatrix}
\]

## 24.

\[T(f(t)) = f''(t)f(t)\]

\[
  \begin{aligned}
    & T(1) = 0(1) = 0
    & T(t) = b(0)(t) = 0\\
    &T(t^2) =c(2)(t^2) = 2ct^2
  \end{aligned}
\]

\[\begin{bmatrix}
  0 & 0 & 0\\
  0 & 0 & 0\\
  0 & 0 & 2
\end{bmatrix}\]

Not isomorphic.

## 27.

\[T(f(2t))\]

\[
  \begin{aligned}
    & T(1) = 1a = 1\\
    & T(t) = b(2t) = 2bt\\
    & T(t^2) = c(4t^2)= 4ct^2
  \end{aligned}
\]

\[\begin{bmatrix}
  1 & 0 & 0\\
  0 & 2 & 0\\
  0 & 0 & 4
\end{bmatrix} \]

Isomorphic.

## 28.

\[T(f(t)) = f(2t) -f(t)\]

Nonisomorhpic; the constant term gets zeroed out, with amtrix 

\[
  \begin{bmatrix}
  0 & 0 &0\\
  0&1 &0\\
  0&  7 & 3
  \end{bmatrix}
\]

## 29.

\[T = f'(t)\]

\[
  \begin{bmatrix}
  0 & 1 & 0\\
  0 & 0 & 2\\
  0 &0 &0
  \end{bmatrix}
\]

Linear, non-isomorphic.

## 31.

\[T = \m{f(0) & f(1)\\
f(2) & f(3)}\]

from $P_2$ to $R^{2 \times 2}$
Substituting, we find

\[T = 
  \begin{bmatrix}
    a & a + b + c\\
    a + 2b + 4c & a + 3b + 9c
  \end{bmatrix}
\]

The matrix should be:

\[
  \begin{bmatrix}
  1 & 0 & 0
  \end{bmatrix}
\]

```{r, results='asis'}
cbind(matrix(rep(rep(c(
  1, 0, 0
), times = 2), times = 3), ncol = 3), sapply(list(c(2, 0, 0), c(5, 0 , 0), c(13, 0, 0)), rep, 2)) %>% 
  mat2latex()
```

## 32.

\[T(f(t)) = f'(t) +t^2=b + 2ct +t^2\]

Nonlinear; that $t^2$ is affine and has no matrix representation.

## 33.

Drop every other term. 
\[T(x_0, x_1, x_2, x_3, x_4, \dots) = (x_0, x_2, x_4, \dots)\]

Linear, since the sums and scalar multiples of parallel terms are the same before and after the transformation, but not isomorphic because there is no way to recover the terms.

## 34.

\[T(x_0, x_1, x_2, \dots) =(0, x_1, x_2, \dots\]

First rule:

\[
  \begin{aligned}
    & T(x + y) = (0, x_0 + y_0, x_1 + y_1)\\
    & T(x) + T(y) = (0, x_0, x_1, \dots) + (0, y_0, y_1, \dots)
  \end{aligned}
\]

Second rule:

\[
  \begin{aligned}
    & kT(x) = (0, kx_1, kx_2, \dots)\\
    & T(kx) = (0, kx_1, kx_2, \dots) = kT(x)
  \end{aligned}
\]

Linear, isomorphic.

## 35.

From $P$ to $V$:
\[T(f(t)) =(f(0),f'(0), f''(0),\dots\]

The first property is satisfied by the sum rule of derivatives.

Second property satisfied:
\[
  \begin{aligned}
    & kT(f(t)) = (kf(0), kf'(0), kf''(0), \dots)\\
    & T(kf(t)) = ((kf(0), kf'(0), kf''(0), \dots))
  \end{aligned}
\]

Not an isomorphism; many functions are undefined at 0. 

## 39.

Let's find the matrix:

\[
  \begin{aligned}
    & T(f) = f''+2f' +f\\
    & = 2c + 2(b + 2ct) + a + bt + ct^2\\
    & = a + 2b + 2c + bt + 4ct + ct^2\\
  \end{aligned}
\]

\[
  \begin{bmatrix}
  1 & 1 & 1\\
  0 & 2 & 4\\
  0 & 0 &1
  \end{bmatrix}
\]

This is clearly invertible, so it's an isomorphism.

## 43.

\[T(f(t) = \begin{bmatrix}
  f(5)\\
  f(7)\\
  f(11)
\end{bmatrix}\]

\[ \begin{bmatrix}
  1 & 5 & 25\\
  1 & 7 & 49\\
  1 & 11 & 121
\end{bmatrix}\]

Non-isomporhic; the constant term gums up the workss.

## 52.

The kernel of $T(M) = M\m{1&2\\3&6}$ is all matrices which have $\m{1\\2}$ in 
their kernel, of the form $v_1 = -2v_2$. 

## 25. 

\[T(f(t)) = f''(t) + 4f'(t)\]

\[
  \begin{aligned}
    & =2c + 4(b + 2ct)\\
    & = 4b + 2c +8ct
  \end{aligned}
\]

\[
  \begin{bmatrix}
  0 & 4 & 2\\
  0 & 0 &8\\
  0 & 0 & 0
  \end{bmatrix}
\]

The kernel is constant functions you idiot.

## 57.

\[T(f(t)) = f'' - 5f' +6f\]

\[
  \begin{aligned}
    & =2c -5(b + 2ct) +6(a + bt + ct^2)\\
    & = 2c - 5b -10ct + 6a + 6bt +6ct^2\\
    & = 6a -5b + 2c +6bt  - 10ct + 6ct^2
  \end{aligned}
\]

The image is $P_2$, the kernel $f(t) = 0$.

## 59.

From $P_2$ to $R^n$
\[
  \begin{aligned}
    & T(f(T)) = f(7)
  \end{aligned}
\]

The matrix:

\[
  \begin{bmatrix}
  1 & 7 &49
  \end{bmatrix}
\]

The matrix kernel makes this clearer, I think.

\[
  \begin{bmatrix}
  -7 & -49\\
  1& 0\\
  0 &1
  \end{bmatrix}
\]

## 66.

For the infinite-dimensional $T(f) = f - f'$, the kernel is $e^x$, the image all other functions.

Working with the subspace of polynomials of degree $\leq n$ such that $f(0) = 0$; no constant term.

## 72.

Is it a subspace? Yes, because $f(0 + g(0)) =0 + 0 =0$, and $kf(0) = 0(0)=0$.The dimension of $Z_n$ is $n$, because a polynomial with no constant term has $n$ terms.

## 75.

\[
  \begin{aligned}
    & T(0_V)=0_W\\
    & 0_V(T) = 0_W\\
    & 0_V= 0_W
  \end{aligned}
\]

\[
  \begin{aligned}
    & T(0_V) = 0_W\\
    & T^{-1}(T(0_V)) = T^{-1}(0_W)\\
    & 0_V = T^{-1}(0_W)
    \end{aligned}
\]

## 76.

We know $T$ is linear, so we need to prove $T^{-1}$ obeys the properties as well.
\[
  \begin{aligned}
    & T^{-1}(f + g)  = T^{-1}(T(T^{-1}(f + g)))\\
    & = T^{-1}(T(T^{-1}(f)) + T(T^{-1}(g)))\\
    & = T^{-1}(f) + T^{-1}(g)
  \end{aligned}
\]

and

\[
  \begin{aligned}
    & kT^{-1}(f) = T^{-1}(kT(T^{-1}(f))\\
    & = T^{-1}(T(k)T^{-1}f))\\
    & = kT^{-1}f
  \end{aligned}
\]
