---
title: "Section 1.4 Problems"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.pos = "", warning = FALSE,
                      tidy = TRUE, fig.align = "center")

library(tidyverse)
library(rlang)
library(matador)
```

## 4.

An $n \times m$ multiplying any vector involves $m^2n$ multiplications. If $B$
is a $m \times p$, then there are $m^2np$.

## 6.

An example of matrices whose products are transposes:

```{r}
A <- square(2, 3, 3, 4)
B <- square((-1)^(as.numeric(A)))

A %*% B
B %*% A
```

# 10.

True or false:

a. Columns 1 and 3 of B the same, columns 1 and 3 of $A$: true

```{r}
square(1, 2, 3, 0, 0, 0, 7, 2 , 8) %*% square(1, 1, 1, 0, 0, 0, 1, 1, 1)
```

b. Rows 1 and 3 of $B$ the same, rows 1 and 3 of $AB$ the same: true

```{r}
square(1, 0, 0, 0, 1, 0, 0, 0, 1) %*% square(1, 1,1, 0, 0, 0, 1, 1, 1, byrow = TRUE)
```

c. Rows 1 and 3 of $A$ the same, so are rows of $AB$.
```{r}
square(1, 2,3, 1, 2, 3, 1, 2, 3) %*% c(0, 1,1)
```

d. $(AB)^2 = A^2B^2: false$

```{r}
A <- square(1, 0, -1, 0)
B <- t(A)

(A %^% 2) %*% (B %^% 2)

(A %*% B) %^% 2
```

## 13. 
Examples of matrices. 

a. $A^2 = -I$. Two 90-degree rotations, with matrix $\m{0 & -\\ 1 & 0}$

b. $B^2=0$: $\m{0 & 2\\0 & 0}$

c.$CD = -DC$: $C = \m{0 & -1\\1 & 0}$, $D = \m{0 & 1\\ 1 & 0}$

d. $EF=0$ $\m{0 & 1\\0 &0}$ and $\m{0&0\\1 & 0}$

```{r}
A <- square(1, -1, 1, 1)
B <- square(1, 1, -1, 1)
A %*% B
B %*% A %*% B
```

## 16.
Since $(AB)x=A(BX)$, then the first column of $AB$ must equal $A$ times the first column of $B$. If it were something else, then $AB=C$ and

\[
  \begin{aligned}
    & (AB)x = Cx \neq A(Bx)
  \end{aligned}
\]

## 18.

Each entry of AB = $\sqrt{n}$

## 39.

$A$ is $3 \times 5$, $B$ is $5 \times 3$, $C$ is $5 \times 1$, $D$ is $3 \times 1$. All entries are 1.

a. $AB$ 3 by 3, all entries 5.

b. $BA$ is $5 \times 5$, all entries 3.

c. $ABD$: $3 \times 1$, all entries 15.

d. $DBA$: undefined

e. $A(B + C)$: undefined

## 40.

How do you get:

a. col 3 of $AB$: each row of $A$'s DP

b. row 1 of $AB$: row 1 of $A$ with each column of $B$

c. $AB_{3,4}$ row 2 of $A$ column 4 of $B$

d. $CDE_{1,1}$: row 1 of $C$ by column 1 of $D$ by column 1 of $E$.

## 41.
THe only matrices for which:

a. $BA = 4A$: $4I$

b. $BA = 1/4I$

c. $BA$ has rows 1 and 3 of $A$ reversed and row 2 unchanged: 

\[\m{0 & 0 & 1\\
0 & 1 & 0\\
1 & 0 & 0}\]

d. All rows of $BA$ are row 1 of $A$: first column of 1s, all others 0.

## 42. 

True or false:

a. If $A^2$ is defined, $A$ must be square.

b. If $AB$ and $BA$ are defined, both are square: false

c. The above, and $AB$ and $BA$ are square: true

d. If $AB=b$ then $A = I$: false, $B$ might be the zero matrix.