---
title: "Chapter 5 Review Problems"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

\newcommand{\rot}[1]{
\begin{bmatrix}
\cos{(#1)} & -\sin{(#1)}\\

\sin{(#1)} & \cos{(#1)}\\
\end{bmatrix}
}
```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  warning = FALSE,
  tidy = TRUE,
  fig.align = "center",
  highlight = TRUE
)

library(tidyverse)
library(rlang)
library(matador)
```
## 1.

\[A = \begin{bmatrix}
  1 & 0\\
  -1 & 1
\end{bmatrix}
 \begin{bmatrix}
   1 & 0\\
   0 & 3
 \end{bmatrix}
 \begin{bmatrix}
   -1 & 0\\
   1 & 1
 \end{bmatrix}\]
 
 \[B = \frac{2}{3} \begin{bmatrix}
   1 & 1\\
   3/2 & 3
 \end{bmatrix}
  \begin{bmatrix}
    2 & 0\\
    0 & 1
  \end{bmatrix}
  \begin{bmatrix}
    3 & -1\\
    -3/2 & 1
  \end{bmatrix}\]
  
## 2.

$\det A = \lambda_1 \lambda_2$; $\det A^{-1} = \frac{1}{\lambda_1 \lambda_2}$

## 3.
The eigenvectors are orthogonal and the values real, a pair of properties unique to symemtric matrices. The trace is 1, the sum of the eigenvalues, the determinant 0, the product. 

I multiply out the diagonalziation to get:
\[ \frac{-1}{5}\begin{bmatrix}
  -4 & 2\\
  2 & -1
\end{bmatrix}\]

## 4.

The eigenvectors remain the same, as will the eigenvalues, because $0^2=0$ and $1^2 =1$.

## 5.

For real $r$, $A +ri$ is invertible
if $A = \m{1 & 2\\-2 & 4}$ because skew-symmetirc matrices have complex eigenvalues, so there will never be a zero eigenvalue. Likewise, adding $cI$ to an invertible Hermitian does not make it noninvertible because Hermitians always have real eigenvalues. 

## 6.

\[A^t = \frac{1}{2} \begin{bmatrix}
  e^{4t} + e^{2t} & e^{4t} - e^{2t}\\
  e^{4t} - e^{2t} & e^{4t} +e^{2t}
\end{bmatrix}\]

From this the solutions for $u_0 = e_1$ and $u_0 = e_2$ just select the columns.

## 7.
If $t$ is 1 year:
\[
  \begin{aligned}
    & P = (1 + \frac{.4}{4})^{4t} =1.1^{4t}\\
    & P =(1 + \frac{1}{2})^t = 1.5^t
  \end{aligned}
\]
The annual interest is somewhat better.

## 8.

a. False. Row exchanges change the eigenvalues.

b. False. $\m{1 & 1\\ 0 & 2}$ is similar to $\m{1 & 0\\0 & 2}$

c. True. If a matrix is unitary and Hermitian, its transpose is real and its
inverse, so $A^2 = I$. If $A^2 = I$ and the matrix is Hermitian, it must be
unitary because $A^{-1} = A$. If unitary and $A^2=I$, the inverse is the Hermitian and also $A$ itself, so $A$ is Hermitian. 

d. False. Consider $A = \m{2&0\\0&1}$, $B = \m{1&2\\0&2}$

## 9.

The bizarro Fibonacci sequence starts with $F_{-1}=1$, so $F_{-2} + 1 =0 \implies F_{-2} = 1$. from there it's the regular sequence: $1,-1, 2,-3, 5, -8$. Matrix I think is $\m{-1 & 1\\0 & 1}$ 

## 10.

## 11.

For any vector $s$ in $S$, $Ps =s$ because $s^{\perp} = 0$, so $Ax = \lambda s \implies \lambda = 1$. For vectors $v$ orthogonal to the subspace, $Pv =0$, again satisfying the criterion. The eigenvalues are $n- \text{rank}(P)$ zeroes and the rest ones.

## 12.

Probably not the intended reasoning, but this could be done by splitting the
column vectors over two matrices, filling with zeroes, and summing.

## 13.

## 14.

\[A = \begin{bmatrix}
  -7 & 20\\
  -4 & 11
\end{bmatrix}\]

\[Au = \begin{bmatrix}
  17\\
  8
\end{bmatrix}\]

\[\begin{aligned}
& u_{k+1} = 1(1^k) \begin{bmatrix}
  5\\
  2
\end{bmatrix} + 2(3^k) \begin{bmatrix}
  2\\
  1
\end{bmatrix}\\
& = \begin{bmatrix}
  5\\
  2
\end{bmatrix} +
3^k \begin{bmatrix}
  4\\
  2
\end{bmatrix}
\end{aligned}\]

## 15.

The eigenvectors, as expected, are complex but orthogonal.
```{r, results='asis'}
A <- matrix(complex(9, imaginary = c(0, 1, 0, -1, 1, -1, 0, 1, 0)), nrow = 3)
mat2latex(round(eigen(A)$vectors, digits = 3))
```

## 16.
From the multiplication

\[ \begin{bmatrix}
  a^2 +bc & b(a+d)\\
  c(a+d) & bc +d^2
\end{bmatrix}\]
Neither $a$ nor $d$ can  be zero, so $bc= - \sqrt a = - \sqrt d$. But then $c$ must be zero for the bottom left to be 0 yet nonzero for the diagonal to be 0. 
If $A = \m{4 & 1\\ 0 &4}$, then a possible square root is $\m{2 & 1/4\\ 0 & 2}$

## 17.
a.

\[du/dt = \frac{1}{8} \Big ( 500 e^t \begin{bmatrix}
  4\\
  1
\end{bmatrix} - 300 e^{-t} \begin{bmatrix}
  4\\
  -1
\end{bmatrix}
\Big )
\]

## 18.

a. True.

b. False. Consider $A = \m{1&0\\0&0}$.

e. False. Consider

\[ \begin{bmatrix}
  1 & 1\\
  0 & 1
\end{bmatrix}\]

d. False. Consider

\[\m{1 & 1\\ 1 & 1}\]

e. False. This is true only of Hermitian matrices.

## 19.

a. The column vectors are linearly independent, and 1 cannot be an eigenvalue, so $K - I$ is invertible.

b. The orthogonal eigenvectors can be converted to a unitary matrix and used to diagonalize $K$.

c. Since $e^{\lambda i}e^{-\lambda i}= e^{\lambda i - \lambda i}=e^0=1$, 
$e^ {\Lambda t}$ is unitary.

d. From the above,
\[
  \begin{aligned}
    & KK^H = U e^{\Lambda t}U^HU (e^\Lambda t)^H U\\
    & = Ue^{\Lambda t} (e^{\Lambda t})^HU^H\\
    & = UU^H\\
    & = I
  \end{aligned}
\]

## 21.

a. \[ M = \begin{bmatrix}
  d & 0 & 0\\
  0 & d^2 & 0\\
  0 & 0 & d^3
\end{bmatrix}\]

b. 
\[M^{-1}AM = \begin{bmatrix}
  1 & d & d^2\\
  1/d & 1 & d\\
  1/d^2 & 1/d & 1
\end{bmatrix}\]

## 22.
The eigenvalues must be all -1, since the standard vectors 
are all flipped. But such a rotation is possible only if 
$\det A =1$ (orientation does not change), and this is impossible if $n$ is odd, since then $\det A = -1$. So $A$ must be a 180-degree rotation matrix, e.g., $\m{-1 & 0 \\0 & -1}$

## 23.

\[
  \begin{aligned}
    & x^tA^T = \lambda_x x^T\\
    & x^TA^TA^Ty = \lambda_y x^TA^Ty\\
    & \lambda_x x^TA^Ty = \lambda_y 
  x^TA^Ty\end{aligned}
\]
Since $\lambda_1 \neq \lambda_2$, this is only true if the eigenvectors are orthogonal.

## 24.

```{r, results='asis'}
A <- sin(1/sqrt(2) *outer(1:3, 1:3) * (pi/4))

S <- eigen(A)$vectors
mat2latex(round(S %*% t(S)))
```

## 25.
\[N = \begin{bmatrix}
  0 & 1\\
  0 & 0
\end{bmatrix}\]
a. $\lambda^3x = N^3x$, and $N^3 =0$

b. Let element $ji=ij$ be nonzero. Then $N^2_{ji}=N^2_{ij}$ would be nonzero as well, so the matrix could never reach 0. This
also holds if any part of the trace is nonzero.

## 26.

a. \[A = \frac{1}{9} \begin{bmatrix}
  4 & 2 & 4\\
  2 & 1 & 2\\
  4 & 2 & 4
\end{bmatrix}\]

b. $\lambda = 1$, since the dimension of the subspace is 1.

c. Since $P^n =P$, we just have to solve $Pu_0$.

\[Pu_k = \begin{bmatrix}
  6\\
  3\\
  6
\end{bmatrix}\]

## 27.

From the trace and determinant rules:

\[A = \begin{bmatrix}
  7 & 6\\
  -25/3 & -7
\end{bmatrix}\]

## 28.

a. Any real $d$ and $c = 5$.

b. $c = 5$ and $d \neq 37/3$

## 29.

In each case the eigenvectors are the columns of $S$ and the eigenvalues are $(2, 1)$ because the matrices are similar.

## 30.

The dominant eigenvector is $ \m{1\\2}$,so

\[ \frac{-a}{3} \begin{bmatrix}
  1\\
  2
\end{bmatrix}\]