---
title: "Chapter 5 True or False"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

```{r}

```

<!--

 <!-- Standard custom LaTeX commands -->
                                                                                                                                                                                                                                                                                                                                                                
                                                                                                                                                                                                                                                                                                                                                                        (toc&d\end{bmatrix}}
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        \newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        \newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
                              \newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.pos = "", warning = FALSE,
                      tidy = TRUE, fig.align = "center")

library(tidyverse)
library(rlang)
library(matador)
```


## 1.

False. A linear transformation that preserves length but not angle is not orthogonal. 

## 2.

True: assuming invertibility, the inverse of the transpose is the transpose of the inverse.

## 3.

Assuming $A$ is square, $A^2$ is also orthogonal because $A_j \cdot A_j = 1$ and $A_j \cdot A_k =0$ by definition.

## 4.

False. $(AB)^T=B^TA^T$ The $2,1$ element in the first case is row 2 of $A$ times column 1 of $B$, in the second column 1 of $A$ times row 2 of $B$.

## 5.

True.

## 6.



## 7.

False. A symmetric matrix may have zeroes on the diagonal, like \[`r mat2latex(square(0, 0, 1, 0, 0, 0, 1,0,0), sink = TRUE)`\]

## 8.

True. If $AA^T=I$, then $A^{-1}=A^T$, which is true only of orthogonal matrices. 

## 9.

False. It is $(x \cdot u)u$

## 10.

True.

## 11.

False. This is true only of parallel vectors.For two opposite-signed vectors, $||x+y||^2 = 0$.

## 12.

True. The determinant of $R$ is the same as that of $A$, and $R$ is identical for the transpose, since $Q^T = Q^{-1}$

## 13.

True. For orthogonal matrices, $A^T=A^{-1}$, and the inverse is also orthogonal.

## 14.

False.

\[
  \begin{aligned}
    & AB = A^TB^T\\
    & = (BA)^T
  \end{aligned}
\]

## 15.

## 16.

False. If $A$ is not square, the projection is $A(A^TA)^{-1}A^T$. If it is square, the projection is $I$.

## 17.

True. If $B$ is symmetric, so is $B^2$, since $B^2=B^TB^T=B^2$.

## 18.

True. 

\[
  \begin{aligned}
    & A^TB^T=B^TA^T\\
    & BA = AB
  \end{aligned}
\]

## 19.

False. The subspaces partition $R^5$ completely, and dimensions can only be integers, so if $V^{\perp}$ has an odd dimension $V^{\parallel}$ has an even one and vice versa.

## 20.

True. This is just a restatement of the $QR$ decomposition.

## 21.

False. Since the columns are all unit length, no scaling of volume occurs. But the matrix may have negative orientation, in which case $\det(A) = -1$.

## 22.

True.

## 23.

True. If this were false, the vector would have greater than unit length.

## 24.

True. By definition, a basis consists of linearly independent vectors, which can always be orthonormalized.

## 25.

False. the columns are not unit length.

## 26.

True.

\[
  \begin{aligned}
    & x = x^{\parallel} + x^{\perp}\\
    & x - x^{\parallel} = x^{\perp}
  \end{aligned}
\]

By definition, this vector is orthogonal to the subspace.

## 27.

False. consider:

\[A = \begin{bmatrix}
  -1/ \sqrt2 & 1/ \sqrt 2\\
  1 / \sqrt2 & 1 / \sqrt 2
\end{bmatrix}\quad B = \begin{bmatrix}
  0 & 1\\
  1 & 0
\end{bmatrix}\]

## 28.

True. Since $A$ is symmetric, the orthogonal complement of the image  is the kernel of the transpose, which is the same as $A$'s kernel since $A^T=A$. So those vectors must be orthogonal.

## 29.

True; proved in an earlier problem.

## 30.

False. This is true of any symmetric matrix, since $A^2=A^2$.

## 31.

## 32.

True. A scaling matrix would not leave some vectors with lengths unchanged.

## 33.

True. If $A^{-1}=A$, then $A_j \cdot A_i=0$ and $A_j \cdot A_j= 1$. 

## 34.

True. If the entries are all positive, then
 they must be in the positive-signed quadrant, which encloses angles of less than 90 degrees in $R^2$ and less in higher dimensions.

## 35.

True. $\ker(B^T) = (B^T)^{\perp} = im(B^T)$

## 36.

True. $A^T$ is always symmetric. 

## 37.

## 38.

True. $\ker(B^T) =  B^{\perp}$, so the image doesn't lose dimension.

## 39.

## 40.

True. This is true only if the matrix is symmetric, which implies $\ker(A^T) = \ker(A)$

## 41.

## 42. 
True. This is a restatement of Cauchy-Swarz.

## 43.

True.

## 44.

True.

## 45.

True. 

## 46.

True. This is the definition of the eigenvalue.

## 47.

False. $R^{2 \times 2}$ has dimension 4. But any orthogonal matrix spans $R^2$, so those basis elements would not be linearly independent and therefore not a valid basis.

## 48.

## 49.

## 50.

True. Projection matrices are always symmetric, which guarantees orthogonal eigenvectors.
