---
title: "Chapter 2.4 Problems"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output:
  pdf_document:
    highlight: "kate"
    df_print: "kable"
    includes:
      in_header: "notes_text_preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  message = FALSE,
  tidy = "styler",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)

library(tidyverse)
```


### 1.

$v_4 = 3v_1 - v_2 + 4v_3$

### 2.

4.

### 3.

In case 1, $v_1 = 0v_2$. In the seond, $v_2 = \frac {b}{a}v_1$. In the third,
$v_3 = \frac {c - be/d}{a}v_1 + e/d v_2$ 

### 4.

Evident form the fact $0/c = 0$ for nonzero $c$ and $0 \pm 0 = 0$.

### 5.

### 6.

The matrix has RREF 

\[
    \begin{bmatrix}
    1 & 0 & -1\\
    0 & 1 & 1\\
    0 & 0 & 0
    \end{bmatrix}
\]


### 7.

The matrix  

\[
    \begin{bmatrix}
    0 & 1 & 1\\
    1 & 0 & 1\\
    1 & 1 & 0
    \end{bmatrix}
\]

reduces to the identity.

### 8.

### 9.


### a.

The biggest possible dimension of a set of vectors in $R^3$ is 3.

### b.

Scalars of each other.

### c.

$0v = 0$ for all vectors $v$.
### 10.

Kernel basis 

\[
    \begin{bmatrix}
    -2 & 3 & 1\\
    1 & 0 & 0\\
    0 & 1 & 0\\
    0 & 0 & 1
    \end{bmatrix}
\]


### 11.

### a.

Line.

### b.

Plane

### c.

Plane

### d.

Not a subspace. 

### 12.

$Ax = b$, $x^TA=c$, false; the zero vector is always in the row space.

### 13.

All 2, as it turns out.

### 14.


### 15.


### 16.

They do.

### 17.

The same way: a row of zeroes.

### 18.



### 19.

$n$, span, greater than or equal

### 20.


### a.

$\begin{bmatrix}1\\1\\1\\1\end{bmatrix}$

### b.

$\begin{bmatrix}c_1\\c_2\\c_3\end{bmatrix}$

### c.

\[
    \begin{bmatrix}
    0 & 0\\
    1 & 1\\
    1 & 0\\
    0 & 1
    \end{bmatrix}
\]


### d.

For the kernel, 

\[
    \begin{bmatrix}
    -1 & 0 & -1\\
    0 & -1 & -\\
    1 & 0 & 0\\
    0 & 1 & 0\\
    0 & 0 & 1
    \end{bmatrix}
\]


### 21.

For the column, $\begin{bmatrix}1 & 1\\-1 & 1\end{bmatrix}$, or a similar basis 
of opposite-sign pairs. FOr the row space, the columns of the transpose themselves
or any scalars of them.

### 22.


### a.

Might not.

### b.

Are not.

### c.

Might not.

### d.

Might not.

### 23.

Then rank is $n$, and at most $m$ and at least 1 if they span $R^m$. If they are a basis, there are $m$ 
linearly independent vectors.


### 24.


### 25.


### 26.

Both true, from the definitions of orthogonal complements.

### 27.

For both the kernel is.

\[
    \begin{bmatrix}
    1\\
    -1\\
    1
    \end{bmatrix}
\]


### 28.


### a.

False, if $m>n$ and rank is $n$

### b.

False, consider $\begin{bmatrix}0 & 1\\0 & 0\end{bmatrix}$

### c.

True.

### d.

False, some vectors may be redundant.

### 29.

At least one is nonzero.

### 30.


### 31.

Say  $W$ has dimension 1 and none of the vectors are a multiple of it.

### 32.

3, 0, 16.

### 33.

$\mathbf{W}^{\perp}$ has dimension 2, which is less than the dimension of $\mathbf{V}$.
Therefor the dimension of vectors in the second subspcae but not the first is 
at least 1.

### 34.

Assume $W ^{\perp} \in V$. That is not a basis for  $V$, leaving at least a 
one-dimensional subspace in $V$ not in the kernel and therefore in $W$.
and exactly  $k$ vectors spanning ensures no redundancy.

### 35.

### a.

False; it has zero or one solutions

### b.

True.

### 36.

6 in each case.

### 37.

The first has dimension $n$ the latter 2 $n(n-1)/2$

### 38.


### a.

All constant functions

### b.

$f(x) = 3x + c$
### c.


### 39.


### 40.


### 41.

Respectively constants, linear functions, quadratics.

### 42.

$1-x, 1 - x^2$

### 43.

Solving for the vector in $R^{3 \times 3}$

```{r}
I <- list(list(1, 0, 0), list(c(0, 1, 0)), list(c(0, 0, 1)))
perms <- sapply(list(c(1, 3, 2), c(2, 1, 3), c(3, 2, 1), c(2, 3, 1), c(3, 1, 2)), \(x) unlist(I[x]))
solve(t(perms) %*% perms) %*% t(perms) %*% unlist(I)
```

### 44.


### 45.

If the expanded matrix is invertible, then $b$  is not a linear combination of 
the preceding columns, meaning no solution exists. But if the expanded matrix is 
singular, then $b$  is such a linear combination, making the equation solvable.
