---
title: "Notes"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')"
output: pdf_document
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.pos = "", warning = FALSE,
                      tidy = TRUE, fig.align = "center")

library(tidyverse)
library(rlang)
library(matador)
```


## 1.

A 90 degree rotation followed by an x-axis projection is:

\[
\m{1 & 0\\
0 & 0}\m{0&1\\0&-1}
\]

\[\m{0 & -1\\
0 & 0}\]


## 4.

\[
  \begin{aligned}
    & z = \frac{y-z}{2}\\
    & Az = A\Big(\frac{y-z}{2}\Big)\\
    & = \frac{1}{2}Ay -  \frac{1}{2}Az
  \end{aligned}
\]


## 6.

3 by 3's for:

a. x-axis projection:

\[
  \begin{bmatrix}
  1 & 0 & 0\\
  0 & 1 & 0\\
  0 & 0 & 0
  \end{bmatrix}
\]

b. $xy$ reflection:

\[
  \begin{bmatrix}
  1 & 0 & 0\\
  0 & 1 & 0\\
  0 & 1 & -1
  \end{bmatrix}
\]

c. $xy$ rotation:

\[
  \begin{bmatrix}
  \sin \theta & -\cos \theta & 0\\
  \cos \theta & \sin \theta & 0\\
  0 & 0 & 1
  \end{bmatrix}
\]

d. Composed 90-degree rotations:

```{r, results = 'asis'}
sin_90 <- 1
cos_90 <- 0
compose_trans(list(square(sin_90, -cos_90, 0, cos_90, sin_90, 0, 0, 0, 1 ),
              square(sin_90, 0, -cos_90, 0, 1, 0, sin_90, 0, cos_90),
              square(1, 0, 0, 0, sin_90, -cos_90, 0, cos_90, sin_90 ))) %>% 
  mat2latex()

```

e. Composed 180-degree:

`
```{r, results = 'asis'}
sin_90 <- sin(pi)
cos_90 <- cos(pi)

compose_trans(list(square(sin_90, -cos_90, 0, cos_90, sin_90, 0, 0, 0, 1 ),
              square(sin_90, 0, -cos_90, 0, 1, 0, sin_90, 0, cos_90),
              square(1, 0, 0, 0, sin_90, -cos_90, 0, cos_90, sin_90 ))) %>% 
  round() %>% 
  mat2latex()
```

## 7.

Find the $P_3$ to $P_4$ matrix of $3t+2$

\[
  \begin{aligned}
    & a = 3t + 2\\
    & b = b(t(3t +2)) = b(3t^2 +2t)\\
    & c = c(t^2(2 + 3t)) = c(3t^3 + 2t^2)\\
    & d = d(t^3(3t+2)) = d(3t^4 + 2t^3)\\
  \end{aligned}
\]

Now the matrix:

\[
  \begin{bmatrix}
  2 & 3 & 0 & 0\\
  0 & 2 & 3 & 0\\
  0 & 0 & 2 & 3\\
  0 & 0 & 0 & 2\\
  0 & 0 & 0  & 3
  \end{bmatrix}
\]

## 12.

$A^{-1}(x + y)= A^{-1}x + A^{-1}y$ by the linearity of the inverse transformation.

## 14.

If $T$ is linear, it has matrix $A$. The $T^2=A(A) =A^2$, which is a composition of linear transformations

## 15. 

What's the transpose matrix for $R^{2 \times 2}$?

\[
  \begin{bmatrix}
  1 & 0 & 0 & 0\\
  0 & 0 &1 & 0\\
  0 & 1 & 0 & 0\\
  0 & 0 & 0 & 1
  \end{bmatrix}
\]

## 17.

a. Matrix of the "right shift" $T(x) = \m{0\\x_1\\x_2\\x_3}$:

\[
  \begin{bmatrix}
   0 & 0 & 0\\
   1 & 0 & 0\\
   0 & 1 & 0\\
   0 & 0 & 1
  \end{bmatrix}
\]

The opposite "left shift":

\[`r mat2latex(cbind(rep(0, 3), diag(nrow = 3)), sink = TRUE)`\]

Naturally, they are left and right inverses of each other.

## 19.

Which nonlinears are invertible?

a. $x^3$: yes

b.$e^x$: no, limited to positive numbers.

c $x + 11$: yes

c.$\cos x$: yes 

## 21.

Why linear transformations can't move the zero vector:

\[T(0 + w) = T(0) + T(w) = T(w)\]

This holds only if $T(0)=0$. Similarly,
$T(0v) = 0T(v)$.

## 23.

If $S$ and $T$ are linear and equal $v$, then $S(T(v)) = S(v) = v$

## 24. 

Say $T(0, v_2) =(0, 0)$. Then $kT(0, v_2)=(0, kv_2) =0  =(0, kv_2)$, but:

\[
  \begin{aligned}
    & T((0, v_2) + w) = (w_1, v_2)\\
    & T((0, v_2)) + T(w) = (0, 0) + w = w
  \end{aligned}
\]

## 25.

Which transformations satisfy which linearity property?

a. $T(v) = v/||v||$: neither

\[
  \begin{aligned}
    & T(v + w) = \frac{v+ w}{||v+w||} \neq v/||v|| + w/||w
    T(cv) = cv /||cv|| \neq cv/ c||v||
  \end{aligned}
\]

b. $T(v) = v_1 + v_2 + v_3$: both

\[
  \begin{aligned}
    & T(v + w) = v_1 + v_2 + v_3 + w_1 + w_2 + w_3 = T(v) + T(w)\\
    &kT(v) = k(v_1 + v_2 v_3) = kv_1 +kv_2 + kv_3
  \end{aligned}
\]

c. $T(v) =(v1, 2v_2, 3v_3)$: obviously a legit diagonal matrix.

d. $T(v) = \max(v)$

Say $\max(w) = k \max(v)$ and the index of $\max(v)$ is not that of $\max(w)$. It satisfies the scalar property only.

\[
  \begin{aligned}
    & T(v+w) = \max(v+w)
    & T(v) + T(w) = \max(v) + k\max(v)
  \end{aligned}
\]

\[
  \begin{aligned}
    & kT(v) = k \max(v)= \max(kv)
  \end{aligned}
\]

## 28.

Find some ranges and kernels:

a. $T(v_1, v_2) = (v_2, v_1)$: $R^2$ and 0.

b. $T(v_1, v_2, v_3) = (v_1, v_2)$ $R^2$ and $\m{0\\0\\1}$

c. $T(v_1, v_2) = (0, 0): $0 and R^2$.

d. $(v_1, v_2) = v_1, v_1$: $(1,0) and (0, 1)$.

## 31. 

A matrix transformation is linear because the dot product, by which multiplication is done, is itself linear.

## 33.

A "universal" transposing matrix cannot exist because we would need to swap the $2,1$ and $1, 2$ elements. But this can only be achieved by a matrix with a a 1 in one of those positions and zeroes elsewhere. Yet the diagonal of the transpose must be $(1,1)$ in order to select columns in the same order. No single matrix can have both these properties. 

Wait a minute, it satisfies the rules of linearity properties but has no matrix!

## 34.

The transpose is involutory ($(A^T) ^T=A$), has the zero kernel, and has a range of all matrices, but $T(M) = -M$ only for diagonal or skew-symmetric matrices.

## 40.

\[\ M = \begin{bmatrix}
  1 & 1\\
  4 & 5
\end{bmatrix}\]

The solution is $\m{5\\-4}$ - the first column of $M^{-1}$, which reverses the effect of the transformation.

## 43. 

The matrix of $T$ is $\Lambda$, since the transformation is just scaling eigenvectors.

## 43.

Because the matrix of a linear transformation is defined as the the transformation applied to the columns of the identity. That means each basis vector has a unique representation in the transformation, which may be inverted.

## 44.

A reflection across the $x$ axis followed by the $y$ axis is also a reflection over $y=-x$

## 45.

$S(T(v)) = \m{y\\ -x}$, also a reflection over the line $y = -x$.

## 49.

If we know $T(V)$ for $n$ distinct vectors, we don't know the transformation for all $v$ unless they're linearly independent.

## 50.

a. A parallelogram

b. Scaling matrices with one value in the trace.

c. Matrices of rank 1.

d. Rotation and reflection matrices.