---
title: "Notes"
author: "Ryan Heslin"
date: "`r Sys.Date()`"
output: pdf_document
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.pos = "", warning = FALSE,
                      tidy = TRUE, fig.align = "center")

library(tidyverse)
library(rlang)
```


```{r}

```

##4.

If $Q_1$ and $Q_2$ are both orthogonal, so is $Q_1Q_2$. 
\[
  \begin{aligned}
    & (Q_1Q_2)(Q_1Q_2)^T=I\\
    & Q_1Q_2Q_2^TQ_1^T = I\\
    & Q_1Q_1^T = I\\
    & I = I
  \end{aligned}
\]

##5.

Show the reflection matrix $R = 2P-I$ is orthogonal. I proved in the last chapter that $R^2=I$, and since it is symmetric, $R^TR=I$, making it orthogonal as well.


## 8.

Example showing that projections onto non-orthogonal vectors sum to more than $b$:


```{r}
b <- 1:2

b - (tcrossprod(c(1, 0)) %*% b) + (tcrossprod(c(1, 2))) %*% b
```

## 9.

If $q_1$ and $q_2$ are orthogonal, the closest combination to $q_3$ is $0q_1 + 0q_2$

## 10.

Given GRaham-Schmidt inputs $a$ and $b$, $a = ||a||u_1||$, and $b = ||a||u_1 + u_1 \cdot b$

## 11. 

Upper triangular orthogonals must be diagonal. Since $Q^2=I$, each diagonal element $Q_{ii}$ must be $\pm 1$. 

## 18.

\[
  \begin{aligned}
    & A = QR\\
    & P = QR((QR)^TQR)^{-1}(QR)^T\\
    & = QR(R^TQ^TQR)^{-1}R^TQ^T\\
    & = QR(R^TR)^{-1}R^TQ^T\\
    & = QQ^T
  \end{aligned}
\]

If $Q$ is square, then the projection is the identity.

## 31.

a. If $Q$ is orthogonal, $Q^{-1}$ is as well by definition.

b. True. For a $3 \times 2$,  $Qx = x_1q_1 + x_2q_2$. Since the columns are unit length, the sum of squares doesn't change.

## 32.

\[x_1 + x_2 + x_3 - x_4 = 0\]

All solutions
\[
  \begin{bmatrix}
  -1 & -1 & 1\\
  1 & 0 & 0\\
  0 & 1 & 0\\
  0 & 0 & 1
  \end{bmatrix}
\]

Non-solutions: $\m{-1\\-1\\-1\\1}$

If $b = (1, 1, 1, 1)$, then its projection $S^{\perp}$ is $` r tcrossprod(1/4 * c(-1, -1, -1, 1)) %*% rep(1, 4)$, leaving $\m{7/8\\7/8\\7/8\\9/8}$ in $S$.