---
title: "Section 9.1 Problems"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
    echo = TRUE,
    comment = "",
    fig.pos = "",
    warning = FALSE,
    tidy = TRUE,
    fig.align = "center",
    highlight = TRUE
)
```

## 1.

$x = 7 3^{5t}$

## 3.

$P=7e^{0.03t}$

## 7.

$x = \frac{-1}{t+1}$, so it approaches 0.

## 9.

$x^{-k}dx = dt$, so $\frac{x^{-k+1}}{-k+1} = t + 1$, implying $x = \frac{1}{((-k+1)(t+1))^{\frac{1}{1-k}}}$.

## 13.

$1990-1778=212$ years elapsed.
Hence the sum is either $P = 450000e^{(1.06)212} = `r 450000 * exp(.06 * 212)`$ or $450000(1.06)^{212} = `r 450000 * (1.06^212)`$.

## 15.

If $k$ is a percentage, the growth rate is $k/100$. So we have

\[
\begin{aligned}
& 2 = e^t(k/100) \\
& \ln 2 = t(k/100) \\
& t = \frac{ln 2}{ k/100}\\
& t = \frac{100 \ln 2}{k}
& t \approx \frac{69}{k}
\end{aligned}
\]


## 22.

By the sum rule of derivatives, $\frac{d(x_1 + x_2)}{dt} = \frac{dx_1}{dt} + \frac{dx_2}{dt}$.
(where $x_1$ and $x_2$ are initial states).
This means the expression equals $Ax_1 + Ax_2 =A(x_1 + x_2)$

## 23.

By linearity, $kx(t) = kSe^{\Lambda t}S^{-1}x_0$

## 24.

The second matrix is diagonal, with one distinct eigenvalue. Adding
a diagonal matrix shifts eigenvalues elementwise but does not change
eigenvectors. Therefore:


\[\begin{aligned}
& c(t) = Se^{(\Lambda + k) t} S^{-1}c_0 \\
& = e^{kt}Se^{\Lambda  t} S^{-1}c_0\\
& e^{kt}x(t) = e^{kt}x(t)
\end{aligned}
\]

## 25.

Again by linearity, $x(kt) = kAx \implies kx(t) = A(kx) \implies x(t) = Ax$.

```{r}
# https://stackoverflow.com/questions/43223579/solve-homogenous-system-ax-0-for-any-m-n-matrix-a-in-r-find-null-space-basi/43225005#43225005
kernel <- function(A) {
    m <- dim(A)[[1]]
    n <- dim(A)[[2]]

    ## QR factorization and rank detection
    QR <- base::qr.default(A)
    r <- QR$rank
    ## cases 2 to 4
    if (r == 0) {
        return(diag(x = 1, nrow = m))
    }
    if (r < min(m, n) || (m < n)) {
        R <- QR$qr[1:r, , drop = FALSE]
        P <- QR$pivot
        F <- R[, (r + 1):n, drop = FALSE]
        I <- diag(1, n - r)
        B <- -base::backsolve(R, F, r)
        Y <- rbind(B, I)
        X <- Y[order(P), , drop = FALSE]
        return(X)
    }
    ## case 1
    matrix(0, nrow = n, ncol = 1)
}
# Solve dx/dt = Ax for a matrix A,
# using standard factorization
solve_differential <- function(A) {
    lambda <- eigen(A, only.values = TRUE)[[1]]
    n <- nrow(A)
    S <- lapply(lambda, \(x) kernel(A - diag(x = x, n))) |>
        do.call(what = cbind)

    if (ncol(S) < n) stop("A cannot be diagonalized without using its Jordan form, everybody panic")
    Lambda <- diag(exp(lambda), nrow = n)
    S_inv <- solve(S)
    list(S = S, Lambda = Lambda, S_inv = S_inv)
}

compute_differential <- function(diff, x0, t) {
    diff$S %*% (diff$Lambda * exp(t)) %*% diff$S_inv %*% x0
}
```


## 42.


The relationship seems to be predator-prey. Only the eigenvectors stabilize in the long run, as is expected. The populations don't make much difference.
```{r}
A <- matrix(c(1.4, 0.8, -1.2, -1.4), nrow = 2)
result <- solve_differential(A)
compute_differential(result, c(3, 1), 10)
```


## 45.


$y$ kill 4 $x$ for every member they lose,
a 4:1 kill ratio.
```{r}
A <- matrix(c(0, -1, -4, 0), nrow = 2)
result <- solve_differential(A)
combos <- expand.grid(x = 1:10, y = 1:10)
results <- combos |>
    as.matrix() |>
    asplit(MARGIN = 1) |>
    sapply(\(x) compute_differential(result, x, 30)) |>
    t()
results <- cbind(combos, results)
```

$x$ seems to require about a 2:1 advantage to prevail.

## 46.

The $x$:$y$ kill ratio is $q/p$. The general solution is:

\[
\frac{1}{2\sqrt{p/q}}\begin{bmatrix} 1 & 1 \\
\sqrt{p/q} & - \sqrt{p/q}
\end{bmatrix}
\begin{bmatrix}
e^{\sqrt{pq}t} & 0 \\
0 & e^{-\sqrt{pq}t}
\end{bmatrix}
\begin{bmatrix}
- \sqrt{p/q} & -1 \\
- \sqrt{p/q} & 1
  \end{bmatrix} x_0
  \]

## 51.

The derivative is a linear transformation, since $\frac{d}{dt}(x(t) + y(t))= \frac{d}{dt}x(t) + \frac{d}{dt}y(t)$
and $\frac{d}{dt}kx(t) = k \frac{d}{dt}x(t)$.

This means $S$ can be brought out of the derivative, since it is invariant with
respect to the derivative, so $\frac{d}{dt}(Sx(t)) = S\frac{d}{dt}x(t)$.


## 52.

All are of the form $e^{\lambda t}x_0$.

## 53.

The solution works out to
\[x(t) = \frac{1}{2i} \begin{bmatrix}
i(e^{(p + qi)t} + e^{(p-qi)t}) \\
e^{(p + qi)t} - e^{(p-qi)t}
\end{bmatrix}\]

which is on the complex plane.
